### 2022-11-16 paper.Rmd
added appendix + page x of y and other stuff

---
title: 'Assignment 2: Interpreting Quantitative Findings'
subtitle: |
  | 
  | University of Glasgow
  | `r source("scripts/utils.R")`
  | Student ID: 2819052
  | Course: Quantitative Methods
  | Number of words: `r RmdWords("paper.Rmd")$num_words`
output:
  pdf_document:
    keep_tex: yes
    latex_engine: pdflatex
    number_sections: yes
  html_document: default
header-includes:
- \usepackage{amsmath} # format math stuff
- \usepackage{booktabs} # makes table nicer
- \usepackage{lastpage} # be able to count total pages
- \usepackage{fancyhdr} # header/footer
- \pagestyle{fancy} # header/footer
- \fancyhead{} # remove header
- \renewcommand{\headrulewidth}{0pt} # remove header
- \fancyfoot[CO,CE]{Page \thepage\ of \pageref{LastPage}} # add Page X of Y
- \usepackage{floatrow} # position tables and figures correctly
- \floatsetup[figure]{capposition=top} # position tables and figures correctly
- \floatsetup[table]{capposition=top} # position tables and figures correctly
bibliography: references.bib # add bibliography
csl: chicago-author-date # choose chicago citation style
link-citations: yes # add links in citations
nocite: '*' # add all used citations
---

```{r setup, include=FALSE}
source("scripts/config.R")
```

\pagebreak
<!-- TC:ignore -->
```{=latex}
\setcounter{tocdepth}{2}
\tableofcontents
```
<!-- TC:endignore -->

\pagebreak

# Introduction
<!-- Word count: approx. 400  -->
<!-- Look at the model in part 2 below. Based on this state a research question and a hypothesis. Remember a hypothesis needs at least two variables. -->

<!-- Word count now: 420  -->

In 1998, the Good Friday/Belfast Agreement was signed and later validated by the voters, and this agreement marked a new era of hope towards peace, equality, and inclusion in Northern Ireland [@galligan2013gender]. The main focus on the agreement and referendum was obviously to stop the violent conflict and seek peace, but one part of the agreement also emphasized an equality agenda [@Hayward2021]. Thus, the agreement also marked the first Northern Irish formal recognition of women's rights to political inclusion [@galligan2013gender]. However, a formal recognition does not necessarily imply that gender equality trickles down into social norms and practices. Therefore, this report examines the contemporary state of women's equality in Northern Ireland.

Perhaps, the most central concept within gender equality is the *gender pay gap*. Disparities in income is an important indicator for gender equality, because it has social, economic, and physiological consequences [@bishu2017gender]. Research on this area have identified several factors that seems to influence a gender pay gap. One such factor can be inequality in access to workplace authority, where women are denied manager or supervisor position although there were equally qualified [@bishu2017gender]. Other factors can be discrimination in hiring or promotion processes, but also lack of gender representation can avoid minorities to even apply for a job or promotion [@bishu2017gender].

In order to examine the current gender equality in Northern Ireland, we therefore focus on the gender pay gap in Northern Ireland. We ask the following research question: *Does gender affect income in contemporary Northern Ireland?* We employ a deductive approach, where we first formulate a hypothesis and subsequently examine it empirically [@bryman2016social 33]. Therefore, our analysis consider the following null hypothesis (H~0~) and alternative hypothesis (H~A~):

* H~0~: Being a woman or man does not affect your income.
* H~A~: Being a woman rather than a man affects a lower income.

The causal relationship that we examine is thus a negative relationship. Although, we call this relationship causal, it is important to clarify that I do not imply that there is anything biological or deterministic that reduce women's income in general. Rather, the causal link is interpreted as a result of the societal discriminatory practices explained in the previous paragraph. To make a convincing inquiry of this causal relationship, we must also control for a number of other factor related to income and gender. This is important to make sure that our analysis do not simply show a spurious correlation as it could in fact be another factor that influences both gender and income. These control variables are introduced in the Data and Methods-section.

<!--  \pagebreak -->
# Data and method
<!-- Word count: approx. 800 -->
<!-- In this section your goal is to demonstrate that you can describe data and that you understand research design and data collection. -->

The research design of this analysis is a cross-sectional design [@bryman2016social 53]. This means that the data for analysis is collected at a single point in time and consists of a sample of respondents from which we seek to infer to a general population of Northern Ireland [@field2012discovering 36]. This section describes this sample and how the data was collected, and then subsequently the operationalized variables to examine our research hypothesis are described.

## Sample and Data Collection

Our sample for the analysis consists of data from Northern Ireland Life and Times Survey (NILT). The data have been collected every year since 1998, but this analysis uses the survey from 2012 [@ark_2012]. The respondents for the NILT survey was chosen from a systematic random sample of addresses. From this sample of 2126 addresses, 1204 questionnaires were fully completed using partly face-to-face- and self-completion questionnaires [@ark_2012]. Our sample of analysis is further reduced as not all respondents have answered our variables of interest, and we are therefore not able to include those respondents in our analysis. In the tables below, we present descriptive statistics for our variables of interest in both the full NILT sample and our sample of analysis. The first tables shows the categorical variables, where a frequency distribution are shown [@fogarty2018quantitative 88]. The second table shows the numerical variables, where we have calculated the mean and standard deviation [@fogarty2018quantitative 95].

<!-- TC:ignore -->
```{r echo = FALSE, results='asis'}
sampling_table_categorical()
sampling_table_numerical()
```
<!-- TC:endignore -->

In the table with categorical variables, we conduct a Chi-squared test to check if there is independence between the samples or not, i.e. whether the frequency distribution is significantly different [@fogarty2018quantitative 176]. For the numerical variables, we conduct an Analysis of Variances (ANOVA) to test whether the means in the two samples are significantly different [@field2012discovering 399ff]. For the categorical variables, we see that for the test of trade union membership, the test value is 9.162, which is significantly different from 0 (with a p-value less than 0.01). This indicates that the frequency distribution in our sample is significantly different from the full NILT sample. There is a larger share (44.6 %) of members of a trade union after removing missing values compared to only 37.3 % in the full NILT sample. Similarly, we see from the ANOVA that the test value (F) for age is significantly different from 0 (with a p-value less than 0.01) meaning that our sample is different from the full NILT sample in the age distribution. We can see that the mean age is only around 46 years compared to 49 years in the full NILT sample. These two findings are important to bear in mind, when we interpret our final conclusion as both age and trade union membership might be related to our dependent variable, income.

The systematic random sampling strategy in the NILT survey is an approach to make the sample representative of our population - in our case all inhabitants of Northern Ireland. It could be relevant to examine the demographic distribution in our sample against the general population of Northern Ireland, but unfortunately that is beyond the scope of this report.

## Dependent, Independent, and Control Variables

Levels of measurement: nominal, ordinal, interval, and ratio [@fogarty2018quantitative 56]

<!-- TC:ignore -->
```{r echo = FALSE, fig.cap="Scatterplot of Income and Sex"}
outcome_var_plot()
```
<!-- TC:endignore -->

To fully examine the relationship between our dependent and independent variable, we employ a multiple linear regression estimated with OLS [fogarty2018quantitative 192ff]. Thus, we are also able to include the control variables to make the acceptance or decline of our hypothesis more convincing.

<!--  \pagebreak -->
# Results and discussion
<!-- Word count: approx. 1300 -->
<!-- In this section your goal is to demonstrate that you can interpret quantitative results. Additional credit is available if you are able to relate these findings to social science theories or if you can put the findings in context. -->

Here comes the regression.

<!-- TC:ignore -->
```{r, echo=FALSE, results='asis'}
source("scripts/regression.R")
```
<!-- TC:endignore -->

Effect sizes [@field2012discovering 57]
Influential data points / outliers [@fogarty2018quantitative 221-222]

## Check for Heteroscedasticity
Here comes a check for heteroscedasticity.

<!-- TC:ignore -->
```{r echo = FALSE, fig.cap="Scatterplot of Fitted Values and Residuals"}
hetero_plot()
#model2 <-coeftest(model, vcov = vcovHC)
```
<!-- TC:endignore -->

## Reliability and Validity

Reliability and validity [@bryman2016social 41]
Internal and external validity [@bryman2016social 41-42]

<!-- \pagebreak -->
# Conclusion
<!-- Word count: approx. 500 -->
<!-- Clearly state your findings. Do the findings raise any questions for future research? -->

\pagebreak

<!-- TC:ignore -->

# References

<div id="refs"></div>

\pagebreak

# Appendix

<!-- TC:ignore -->
```{r echo = FALSE, fig.cap="Barplot of Sex"}
control_vars_plot(nilt$rsex, label='Sex')
```
<!-- TC:endignore -->

<!-- TC:ignore -->
```{r echo = FALSE, fig.cap="Barplot of Religion"}
control_vars_plot(nilt$religcat, label='Religion')
```
<!-- TC:endignore -->

<!-- TC:ignore -->
```{r echo = FALSE, fig.cap="Barplot of Sexual Orientation"}
control_vars_plot(nilt$orient, label='Sexual Orientation')
```
<!-- TC:endignore -->

<!-- TC:ignore -->
```{r echo = FALSE, fig.cap="Barplot of Constitutional View"}
control_vars_plot(nilt$uninatid, label='Constitutional View')
```
<!-- TC:endignore -->

<!-- TC:ignore -->
```{r echo = FALSE, fig.cap="Barplot of Trade union membership"}
control_vars_plot(nilt$tunionsa, label='Trade union membership')
```
<!-- TC:endignore -->

<!-- TC:ignore -->
```{r echo = FALSE, fig.cap="Barplot of Supervisor"}
control_vars_plot(nilt$rsuper, label='Supervisor')
```
<!-- TC:endignore -->

<!-- TC:ignore -->
```{r echo = FALSE, fig.cap="Density plot of Age"}
numeric_vars_plot(nilt$rage, label='Age')
```
<!-- TC:endignore -->

<!-- TC:ignore -->
```{r echo = FALSE, fig.cap="Density plot of Income"}
numeric_vars_plot(nilt$persinc2, label='Income')
```
<!-- TC:endignore -->